Improving the Image Acquisition Rate \\ of an Atomic Force Microscope \\ Through Sub-sampling and Reconstruction

### 2019-06-28 latex/afm-cs-main.tex
Fix final typos, adjust RDI definition.

Our previous work on the $\mu$-path pattern, using theoretical
calculations, simulations, and a preliminary implementation,
demonstrates that the approach can achieve scanning time
reduction while maintaining faithful image reconstruction
\cite{maxwell2014compressed,Luo:2015tu, braker_hardware_2018}. 
The main contributions of this paper relative to that past work are as follows.
% \begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\begin{itemize}
\item We introduce a new reconstruction algorithm which is designed to
minimize the artifacts arising from the structure of the $\mu$-path
scanning pattern. This algorithm is described and compared to existing
reconstruction methods through simulations in Sec. \ref{sec:reconstructionMethods}.
\item In our prior experimental prototype \cite{braker_hardware_2018}, we used very basic integral controllers for all three axes for both raster and $\mu$-path scans. Here, we design compensators for each axis using standard loop-shaping techniques (Sec.~\ref{sec:control}), which result in a substantial increase in closed-loop bandwidth (e.g., 50 Hz to 450~Hz for the $Z$-axis).
We show in Sec.~\ref{sec:ff_control} that one of the limitations of $\mu$-path scanning is the discontinuous $X$-axis reference (e.g., in moving between $\mu$-paths) which excites the cross-coupling modes between the $X$ and $Z$ axes, an effect which neither our prior experimental work nor simulation studies accounted for. We mitigate the cross coupling with a feed-forward design.
\item We experimentally compare (Sec. \ref{sec:results}) $\mu$-path scanning to not only full density raster scans but also to coarse raster scans which are interpolated to a full image. This is in contrast to typical sub-line sampling which takes scan lines at random. The clear advantage to a coarse raster scan is that it requires no modification to the instrument, only to the post-processing.
\item The bulk of this work is based on experiment, not simulation, in contrast to prior work by ourselves \cite{luo2015comparison,maxwell2014compressed} and others \cite{han_optimal_2018,jensen_reconstruction_2013,oxvig_structure_2017}. Our main focus here is on delineating the practical issues which must be addressed to effectively realize $\mu$-path scanning, which has only seen limited attention in the past.
% \end{enumerate}
\end{itemize}

@@

A simulation comparison between BPDN and BPVV is shown in
Fig.~\ref{fig:BPTV_demonstration}. The first column shows two raster
scans: the top is a 512 by 512 pixel image of a CS-20NG grating (Ted Pella, USA) acquired
with an Agilent 5400 AFM and the bottom is a 256 by 256 pixel image of
DNA acquired with an Agilent 5500 AFM. For the CS-20NG grating, we
sampled 15\% of the pixels with 50 pixel $\mu$-paths and 25\%
of the DNA image with 25 pixel long $\mu$-paths. The second columns
shows the full frame reconstructions using BPDN and the third columns
shows the result using BPVV (with $\alpha = 0.75$ and $\beta = 0.1$). The fourth column zooms in on the region
indicated by the red box in column one. Similarly, the fifth and six
columns show the detail of the reconstructions. The reconstruction
results show the artifacts apparent in the BPDN reconstruction are
largely mitigated using BPVV. Thus, in the remainder of the paper, we
focus exclusively on BPVV. These reconstructions wth BPVV took about
1.5 seconds for the DNA image and about 7 seconds for the
CS-20NG (on a typical laptop computer).

@@

Transition between, and operation in, these different stages is
implemented as a simple state machine, summarized in
Table~\ref{tab:cs_tasks}. The main job of the state machine is to
adjust the reference signals to accomplish each task. In the
following, $r_X$ and $r_Y$, $r_Z$ refer to the references for the
$X$, $Y$, and $Z$ axes, respectively. The $Z$-axis setpoint $r_Z$
takes on values $r_{Z,s}$ (the scanning setpoint) and $\rzup$ (the
retraction setpoint).

@@

In order to minimize the scanning time, tip tip should be moved as fast as possible when trainsitioning between $\mu$-path locations (state (1)). Since such speeds can be beyond the bandwidth of the $z-$axis controller, it is important to reduce the tip-sample interaction force to prevent tip wear and sample damage. In our initial implementation \cite{braker_hardware_2018}, 
%we insisted that the during
this was achieved by insisting that during the
tip-retraction (state (5)), the cantilever tip should fully break
contact with the surface. Here, we impose no such requirement. Rather,
we change the $Z$-setpoint during tip-retraction to only pull away far
enough that the deflection signal stays low (i.e., below the scanning
setpoint), even while we run across the surface. We choose \rzup so
that the interaction between the tip and sample becomes purely
adhesive. In general, stable imaging is not possible with the setpoint at \rzup and
complete dis-engagement occasionally occurs. This can be seen in the bottom pane of
Fig.~\ref{fig:cs_cycle} during the second $XY$-move (black portion) where the tip detaches partway through the move.
%Fig.~\ref{fig:afm_bd_dinv} which selects either the scanning setpoint $r_{Z,s}$ or the retraction setpoint $r_{Z,\textrm{up}}$ based the current state.
When dis-engagement does occur, the $Z$-axis control loop implements an anti-windup scheme to prevent the control from saturating.

Only moving the tip into the adhesive region during the $XY$ move saves considerable time
compared to completely detaching the tip between each $\mu$-path. By essentially eliminating the normal force into the sample, we are able to take dozens of images with a single probe without a noticeable degradation in image quality and is a key component to the new quality metric we introduce in Sec.~\ref{sec:rdi}. However, completely characterizing the damage done to the tip and sample with this scheme compared complete detachment (which can also damage the probe \cite{meyer_atomic_1992}), remains an outstanding goal.

@@

\subsection{Limitations of SSIM and PSNR}\label{sec:limits}
The SSIM and PSNR have become reasonably popular in studies comparing
simulated AFM image reconstructions \cite{oxvig_structure_2017,
  han_optimal_2018, Luo_nano_2015}. However, it remains somewhat of an
open question how to best compare experimental AFM images. While we
use the SSIM and PSNR in Sec.~\ref{sec:results}, we believe those
numbers should be interpreted with some caution. In our experimental
setup, actuation in the $XY$-plane happens via the nPoint piezo stage
and the $XY$-direction of the original piezo scanner is uncontrolled.
This leads to substantial drift between images. To illustrate this, we
took a sequence of 6 raster scans of a sample grating, each at a 1~Hz
scan rate over a 5 micron by 5 micron area. The left image of
Fig.~\ref{fig:baseline_errors} shows the error between the first and
sixth image. The PSNR and SSIM between the first and second images are
\baseline{one2twoNoAlgnPsnr} and \baseline{one2twoNoAlgnSsim}; between
the first and sixth are \baseline{one2sixNoAlgnPsnr} and
\baseline{one2sixNoAlgnSsim}, respectively, even though on their own
each image appears to have good quality.

@@

\subsection{A new metric}
The PSNR and SSIM metrics only assess the similarity of two images.
However, in AFM imaging, pure image quality is not the only concern,
particularly for delicate samples. While damage to the specimen is not really a concern while imaging a
hard calibration grating, it plays a prominent role for biological
samples~\cite{ando_highspeed_2008}, though damage to the probe itself
is still a concern even with hard samples. In general, the faster the scan rate, the more
difficulty the $Z$-axis control will have in following the sample
surface, which results in a larger $Z_d$ over more time. The actual
damage done by a given $Z_d$ will depend on the spring constant of the
cantilever and the softness of the specimen. To capture these concerns, we propose a
metric, which we call the Relative Damage Index (RDI) that computes
the power in the deflection signal for a given scan. The RDI,given by

\begin{align}
  \text{RDI} &= \frac{1}{T_s}\sum_{k=1}^N \frac{1}{k} \left(Z_{d,k} - r_{Z,s}\right)^2, \label{eqn:RDI}
\end{align}
accumulates the deflection signal as a measure of the net force applied to the specimen  over time.

>>>>

Our previous work on the $\mu$-path pattern, using theoretical
calculations, simulations, and a preliminary implementation,
demonstrates that the approach can achieve scanning time
reduction while maintaining faithful image reconstruction
\cite{maxwell2014compressed,Luo:2015tu, braker_hardware_2018}. 
The main contributions of this paper relative to that past work are as follows.
% \begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\begin{itemize}
\item We introduce a new reconstruction algorithm which is designed to
minimize the artifacts arising from the structure of the $\mu$-path
scanning pattern. This algorithm is described and compared to existing
reconstruction methods through simulations in Sec. \ref{sec:reconstructionMethods}.
\item In our prior experimental prototype \cite{braker_hardware_2018}, we used very basic integral controllers for all three axes for both raster and $\mu$-path scans. Here, we design compensators for each axis using standard loop-shaping techniques (Sec.~\ref{sec:control}), which result in a substantial increase in closed-loop bandwidth (e.g., 50 Hz to 450~Hz for the $Z$-axis).
We show in Sec.~\ref{sec:ff_control} that one of the limitations of $\mu$-path scanning is the discontinuous $X$-axis reference (e.g., in moving between $\mu$-paths) which excites the cross-coupling modes between the $X$ and $Z$ axes, an effect which neither our prior experimental work nor simulation studies accounted for. We mitigate the cross coupling with a feed-forward design.
\item We experimentally compare (Sec. \ref{sec:results}) $\mu$-path scanning to not only full density raster scans but also to coarse raster scans which are interpolated to a full image. This is in contrast to typical sub-line sampling which takes scan lines at random. The clear advantage to a coarse raster scan is that its implementation only requires adjustments to the data collection and post-processing. 
  % it requires no modification to the instrument, only to the post-processing.
\item The bulk of this work is based on experiment, not simulation, in contrast to prior work by ourselves \cite{luo2015comparison,maxwell2014compressed} and others \cite{han_optimal_2018,jensen_reconstruction_2013,oxvig_structure_2017}. Our main focus here is on delineating the practical issues which must be addressed to effectively realize $\mu$-path scanning, which has only seen limited attention in the past.
% \end{enumerate}
\end{itemize}

@@

A simulation comparison between BPDN and BPVV is shown in
Fig.~\ref{fig:BPTV_demonstration}. The first column shows two raster
scans: the top is a 512 by 512 pixel image of a CS-20NG grating (Ted Pella, USA) acquired
with an Agilent 5400 AFM and the bottom is a 256 by 256 pixel image of
DNA acquired with an Agilent 5500 AFM. For the CS-20NG grating, we
sampled 15\% of the pixels with 50 pixel $\mu$-paths and 25\%
of the DNA image with 25 pixel long $\mu$-paths. The second columns
shows the full frame reconstructions using BPDN and the third columns
shows the result using BPVV (with $\alpha = 0.75$ and $\beta = 0.1$). The fourth column zooms in on the region
indicated by the red box in column one. Similarly, the fifth and six
columns show the detail of the reconstructions. The reconstruction
results show the artifacts apparent in the BPDN reconstruction are
largely mitigated using BPVV. Thus, in the remainder of the paper, we
focus exclusively on BPVV. These reconstructions wth BPVV took about
1.5 seconds for the DNA image and about 7 seconds for the
CS-20NG (on a typical desktop computer).

@@

Transition between, and operation in, these different stages is
implemented as a simple state machine, summarized in
Table~\ref{tab:cs_tasks}. The main job of the state machine is to
adjust the reference signals to accomplish each task. In the
following, $r_X$, $r_Y$, and $r_Z$ refer to the references for the
$X$, $Y$, and $Z$ axes, respectively. The $Z$-axis setpoint $r_Z$
takes on values $r_{Z,s}$ (the scanning setpoint) and $\rzup$ (the
retraction setpoint).

@@

In order to minimize the scanning time, the tip should be moved as fast as possible when trainsitioning between $\mu$-path locations (state (1)). Since such speeds can be beyond the bandwidth of the $z-$axis controller, it is important to reduce the tip-sample interaction force to prevent tip wear and sample damage. In our initial implementation \cite{braker_hardware_2018}, 
%we insisted that the during
this was achieved by insisting that during the
tip-retraction (state (5)), the cantilever tip should fully break
contact with the surface. Here, we impose no such requirement. Rather,
we change the $Z$-setpoint during tip-retraction to only pull away far
enough that the deflection signal stays low (i.e., below the scanning
setpoint), even while we run across the surface. We choose \rzup so
that the interaction between the tip and sample becomes purely
adhesive. In general, stable imaging is not possible with the setpoint at \rzup and
complete dis-engagement occasionally occurs. This can be seen in the bottom pane of
Fig.~\ref{fig:cs_cycle} during the second $XY$-move (black portion) where the tip detaches partway through the move.
%Fig.~\ref{fig:afm_bd_dinv} which selects either the scanning setpoint $r_{Z,s}$ or the retraction setpoint $r_{Z,\textrm{up}}$ based the current state.
When dis-engagement does occur, the $Z$-axis control loop implements an anti-windup scheme to prevent the control from saturating.

Only moving the tip into the adhesive region during the $XY$ move saves considerable time
compared to completely detaching the tip between each $\mu$-path. By essentially eliminating the normal force into the sample, we are able to take dozens of images with a single probe without a noticeable degradation in image quality and is a key component to the new quality metric we introduce in Sec.~\ref{sec:rdi}. However, completely characterizing the damage done to the tip and sample with this scheme compared to complete detachment (which can also damage the probe \cite{meyer_atomic_1992}), remains an outstanding goal.

@@

\subsection{Limitations of SSIM and PSNR}\label{sec:limits}
The SSIM and PSNR have become reasonably popular in studies comparing
simulated AFM image reconstructions \cite{oxvig_structure_2017,
  han_optimal_2018, Luo_nano_2015}. However, it remains somewhat of an
open question how to best compare experimental AFM images. While we
use the SSIM and PSNR in Sec.~\ref{sec:results}, we believe those
numbers should be interpreted with some caution. In our experimental
setup, actuation in the $XY$-plane happens via the nPoint piezo stage
and the $XY$-direction of the original piezo tube scanner is uncontrolled.
This leads to substantial drift between images. To illustrate this, we
took a sequence of 6 raster scans of a sample grating, each at a 1~Hz
scan rate over a 5 micron by 5 micron area. The left image of
Fig.~\ref{fig:baseline_errors} shows the error between the first and
sixth image. The PSNR and SSIM between the first and second images are
\baseline{one2twoNoAlgnPsnr} and \baseline{one2twoNoAlgnSsim}; between
the first and sixth are \baseline{one2sixNoAlgnPsnr} and
\baseline{one2sixNoAlgnSsim}, respectively, even though on their own
each image appears to have good quality.

@@

\subsection{A new metric}
The PSNR and SSIM metrics only assess the similarity of two images.
However, in AFM imaging, pure image quality is not the only concern,
particularly for delicate samples. While damage to the specimen is not really a concern while imaging a
hard calibration grating, it plays a prominent role for biological
samples~\cite{ando_highspeed_2008}, though damage to the probe itself
is still a concern even with hard samples. In general, the faster the scan rate, the more
difficulty the $Z$-axis control will have in following the sample
surface, which results in a larger deflection signal $Z_d$ over more time. The actual
damage done by a given $Z_d$ will depend on the spring constant of the
cantilever and the softness of the specimen. To capture these concerns, we propose a
metric, which we call the Relative Damage Index (RDI) that computes
the power in the deflection signal for a given scan. The RDI, given by

\begin{align}
  \text{RDI} &= \frac{1}{T_s}\sum_{k\in I} \frac{1}{k} \left(Z_{d,k} - r_{Z,s}\right)^2, \label{eqn:RDI}\\
  I & = \{k : r_{Z,k} = r_{z,s}\}
\end{align}
accumulates the deflection signal while scanning as a measure of the net force applied to the specimen  over time.
